---
title: "Code Demo - Adapted from UChicago ECMA3132 Final Project"
author: "Raymon Yue"
date: "7/30/2022"
output: html_document
---
In this Econometrics class at UChicago, I chose to write my final paper on the effects of mandatory paid maternity leave on women's wages. 
I decided to use data from the Bureau of Labor Statistics, with by quarterly, by industry, by county average wage estimates 
based on quarterly surveys conducted by the government. This problem is interesting because intuitively, paid maternity leave laws 
could postively impact women's wages because it gives child rearing mothers job safety, encouraging them to stay at the same work place
and accumulate more work experience and reach higher positions. However, it could also negatively impact women's wages, as employers may
intentionally lower women's wages so they don't bear all the costs of the the paid maternity leave policy. U.S. is a particularly interesting
place to study this, as it is one of the last countries in the world without national paid maternity leave laws, and state legislatures have
to pass these laws on their own and at their own pace. The focus of this research is on New York (which enacted paid maternity leave in 2018),
and its Northeastern U.S. neighbors, some of which passed similar laws much earlier (e.g. NJ, 2013), while others have yet to passed them (e.g. PA).

In this Rmd notebook, I gather publicly available data on BLS.gov, clean them, and perform diff-in-diff regressions on it.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
```

A list of industries where women make up more than two thirds of the employees. Their average wages are used as a proxy for women's wages
```{r}
industry<-c(51912,5614,6111,62221,6211,6212,62131,62132,
             6213,6214,6216,623,6239,45439,54194,6241,62421,
              6244,612112,612113,8129,8134)
```

In the code below, I use nested loops to download many different data files and retrieve the information I need and turn them into a single data frame.
```{r}
# a seperately downloaded dictionary of the fips code of every U.S. county
counties<-read_csv("/Users/RaymonYue/Desktop/fips.csv")$fips
# I declared an empty data frame here so I can used rbind in the loop to add rows to it
ne_data <- data.frame(matrix(ncol = 12, nrow = 0))
colnames(ne_data) <- c("area_fips","own_code","industry_code","year","qtr","qtrly_estabs","month1_emplvl",
                            "month2_emplvl","month3_emplvl","avg_wkly_wage","oty_avg_wkly_wage_chg","oty_avg_wkly_wage_pct_chg")
# these are the first two digits of the FIPs code of states in northestern U.S., the region covered by the sample
# try areas <-c(36,42) for a quicker trial run
areas <-c(9,10,23,25,33,34,36,42,44,50)s
```

```{r}
for(a in counties){
  #checks if a county is in the states we sampled
  if(as.integer(a/1000) %in% areas){
    # the state of Connecticut has FIPs starting with 09, 
    # so special attention is needed to make sure the FIPs code string is 5 digits  
    if(as.integer(a/1000)<10){
      astr<-paste("0",a,sep="")
    }else{
      astr<-toString(a)
    }
    # next, I go through every year in the selected sample, 
    # though the study focuses on 2016-2020, I grabbed extra data
    # because DID method requires checking parallel trends assumption before the studied period
    # replace with for(y in 2016:2021) for quicker run
    for(y in 2014:2021){
      for(q in 1:4){
        # the data for 2021 Q4 has not been made public yet at the time of my study, 
        # so I have to make accomodations for it in my loop and skip it
        if(y == 2021 && q==4){
          break
       }
        url<-paste("http://www.bls.gov/cew/data/api/",y,"/",q,"/","area","/",astr,".csv",sep="")
        download.file(url,"mydata.csv")
        temp<-read.csv("mydata.csv")
        file.remove("mydata.csv")
        temp<-subset(temp,select=c(area_fips,own_code,industry_code,year,qtr,qtrly_estabs,month1_emplvl,
                            month2_emplvl,month3_emplvl,avg_wkly_wage,oty_avg_wkly_wage_chg,oty_avg_wkly_wage_pct_chg))
        temp<-filter(temp,own_code==5)
        ne_data <-rbind(ne_data ,filter(temp, industry_code %in% industry))
    }
  }
  next
  }
}
```
Doing an extra step to find the average quarterly employment level, used for the weighted average calculations later
```{r}
ne_data<-ne_data%>%mutate(q_emplvl = (month1_emplvl+month2_emplvl+month3_emplvl)/3)
```
Now I further clean the data, limiting it only to the states of NY and PA, and to the 2016-2020 time period, for a simple DID regression analysis. Since the law went into effect in NY since January 2018, >2017 will be treated time and NY will be the treated state
```{r}
basic_nypa<-ne_data%>%
  filter(year %in% (2016:2020))%>%
  filter(avg_wkly_wage>0)%>%
  filter(as.integer(area_fips/1000) %in% c(36,42))%>%
  filter(as.integer(area_fips%%1000)!=0)%>%
  mutate(treat_time = year>2017)%>%
  mutate(treat_state = as.integer(area_fips/1000)==36)%>%
  mutate(DID = treat_state*treat_time)
```
The regression model includes all three dummies, most importantly the DID estimator which is the interaction term of treated 
time and state, and industry fixed effects
```{r}
summary(lm(avg_wkly_wage~treat_time+treat_state+DID+factor(industry_code),data=basic_nypa))
```

Here I construct an alternative data set where the average weekly wages are weighted averages across industries, with average 
quarterly employment level as the weight
```{r}
wa_wage_nypa<-basic_nypa%>%
  subset(select = c(area_fips, year, qtr, avg_wkly_wage,q_emplvl))%>%
  mutate(avg_wkly_wage =  avg_wkly_wage*q_emplvl)

wa_wage_nypa<-subset(aggregate(wa_wage_nypa, 
                               by=list(wa_wage_nypa$area_fips, 
                                       wa_wage_nypa$year, 
                                       wa_wage_nypa$qtr), sum),
                     select=c(Group.1, Group.2, Group.3, avg_wkly_wage,q_emplvl))

wa_wage_nypa<-wa_wage_nypa%>%
  mutate(avg_wkly_wage = avg_wkly_wage/q_emplvl)%>%
  rename(area_fips = Group.1, year = Group.2 , qtr = Group.3)%>%
  mutate(treat_time = year>2017)%>%
  mutate(treat_state = as.integer(area_fips/1000)==36)%>%s
  mutate(DID = treat_state*treat_time)
  
```
And once again performing the DID regression, this time with the weighted average weekly wage, without industry fixed effect
```{r}
summary(lm(avg_wkly_wage~treat_time+treat_state+DID,data=wa_wage_nypa))
```


